Block bootstrap consistency\\under weak assumptions

### 2016-09-27 Paper.tex
[#27] Make very small grammar fixes in paper

To see when Theorem~\ref{T1}'s Assumption~2 is satisfied, consider two
other parametrizations of this DGP. First, let $\Delta_n = \Delta_0$
and $T_n = T_0$ for some finite $T_0$. This mimics a setting where a
researcher intends to apply the bootstrap to a post-break sample, but
mistakenly includes a small number of pre-break observations as
well. In this case, Assumption 2 holds hold as long as
$M_n = o(\sqrt{n})$.  Similarly, suppose
$\Delta_n = \Delta / \sqrt{n}$ and $T_n = \pi n$, which mimics a
setting with a moderate break with unknown timing.  Then
$D_n \approx \Delta c (1 - c)$ and Assumption 2 again holds as long as
$M_n = o(\sqrt{n})$. Consequently, using block sizes of length
$o(\sqrt{n})$ ensures that Theorem 1's assumptions on the dispersion
of the $\mu_{nt}$ hold when there is a moderate amount of
heterogeneity---either because the sample is chosen to lie after a
large break estimated break or because any breaks are smaller in
magnitude and, consequently, potentially difficult to detect. When the
degree of heterogeneity is very small or nonexistent, so
$\sum_t (\mu_{nt} - \bar \mu_n)^2$ grows even slower, then larger
block lengths can also satisfy this condition.

@@

\begin{lem}\label{L3}
Under the conditions of Theorem~\ref{T1},
\begin{equation}\label{eq:40}
  \E \Big(\sum_{j=1}^{J_n} \Big( \sum_{t = K_{n,j-1}+1}^{K_{nj}}
  (\mu_{nt}^{*} - \bar \mu_n)\Big)^2\Big) = o(n p_n)
\end{equation}
and $\bar \mu_n^{*} = \E^{*} \bar \mu_n^{*} + o_{L_2}(n^{-1/2})$.
\end{lem}
\begin{proof}
  The second result, on $\bar\mu_n^*$ is an immediate implication of
  Equation~\eqref{eq:40}. To show~\eqref{eq:40}, observe that
  \begin{align*}
    \E \Big(\sum_{j=1}^{J_n} &\Big( \sum_{t= K_{n,j-1} +1}^{K_{nj}}
    (\mu_{nt}^{*} - \bar \mu_n)\Big)^2\Big)= \\
    &= \E\Big(\sum_{j=1}^{J_n} \big\lvert \sum_{s,t = K_{n,j-1}+1}^{K_{nj}}
    \E \big((\mu_{ns}^{*} - \bar{\mu}_n)
    (\mu_{nt}^{*} - \bar{\mu}_n) \mid \Ms \big) \big\rvert \Big)\\
    &\leq \E\Big( \sum_{j=1}^{J_n}
    \sum_{k,l = -n}^{n} \tfrac{1}{n} \sum_{\tau=0}^{n-1} \big\lvert
    (\mu_{n,\tau+k} - \bar{\mu}_n) (\mu_{n,\tau+l} - \bar{\mu}_n) \;
    1\{ \tau + k, \tau + l \in I_n(\tau, M_{nj})\} \big\rvert \Big)\\
    &\leq \E\Big(\sum_{j=1}^{J_n} \Big[
    \sum_{k = -n}^{n} \big(\tfrac{1}{n} \sum_{\tau=0}^{n-1}
    (\mu_{n,\tau+k} - \bar{\mu}_n)^2
    1\{ \tau + k \in I_n(\tau, M_{nj})\}\big)^{1/2}\Big]^{2}\Big)\\
    &= \E\Big(\sum_{j=1}^{J_n}
    M_{nj}^2 \Big\lvert \tfrac{1}{n} \sum_{\tau=0}^{n-1}
    (\mu_{n,\tau} - \bar{\mu}_n)^2 \Big\lvert\Big)\\
    &= o(n p_n).
  \end{align*}
  The first equality follows from the LIE. The first inequality
  exploits the fact that the conditional expectation in the previous
  expression averages over the block's starting period but treats
  every other random variable as contstant. (The argument of the
  indicator function and the indices of the summations are
  deliberately chosen to accomodate blocks that include observations
  from both the beginning and the end of the series.) The second
  inequality is an application of the Cauchy-Schwarz inequality to the
  inner-most summation. The next equality holds because the
  summation
  \[
    \sum_{\tau=0}^{n-1} (\mu_{n,\tau+k} - \bar{\mu}_n)^2
    1\{ \tau + k \in I_n(\tau, M_{nj})\}
  \]
  is zero for all but $M_{nj}$ values of $k$ and is a nonzero constant
  for those $k$. The last line holds by Lemma~\ref{L1} and assumption.
\end{proof}

@@

\begin{proof}
  We will only present a proof of~\eqref{eq:30}. Uniform integrability
  of the first family of random variables is simply a restatement of
  that expression and unifom integrability of $Z_{nj}^{*2} n / M_{nj}$
  follows an abbreviated version of the proof of~\eqref{eq:30}.

@@

  Since $Z_{nj}^*$ is a martingale difference array, Doob's maximal
  inequality for martingales \citep[see][Theorem 15.15, for
  example]{Dav:94} along with the construction of $Z_{nj}^{*}$ gives
  a bound for the first term on the \allcaps{RHS} of~\eqref{eq:6},
  \begin{align*}
  \E \Em^* & \Big[\ind(A) \tfrac{n}{m'}\max_{m = 1,\dots, m'}
  \Big(Z_n^*(\tau, K_{n,J(\tau)} - \tau)
  + \sum_{j=J(\tau)+1}^{J(\tau + m)-1} Z_{nj}^*\Big)^2\Big] \\
  &\leq \tfrac{1}{m'} \sum_{u = 0}^{n-1} \E \Big[\Em \big[
  \ind(A) Z_n(u, K_{n,J(\tau)} - \tau)^2\big] +
  \sum_{j=J(\tau)+1}^{J(\tau + m')-1} \Em \big[\ind(A) Z_n(u, M_{nj})^2]\Big] \\
  &\leq \tfrac{3}{m'} \sum_{u = 0}^{n-1} \E\Big[\Em\big[
  \ind(A) Z'_n(u, K_{n,J(\tau)} - \tau)^2\big]
  + \tfrac{1}{n} \Pm[A] \Big(\sum_{t \in I_n(u, K_{n,J(\tau)} - \tau)}
  (\mu_{nt} - \bar \mu_n)\Big)^2 \\
  & \quad+ \sum_{j=J(\tau)+1}^{J(\tau + m')-1}
  \Em \big[\ind(A) Z_n'(u, M_{nj})^2]
  + \tfrac{1}{n} \Pm[A]
  \sum_{j=J(\tau)+1}^{J(\tau + m')-1}\Big( \sum_{t \in I_n(u,M_{nj})}
  (\mu_{nt} - \bar{\mu}_n) \Big)^2 \\
  & \quad+ \Em\big( \ind(A) Z_n'(u, n)^2 \big)
  \sum_{j=J(\tau)}^{J(\tau+m')-1} M_{nj}^2/n^2 \Big] \\
  &\leq (3\delta/m') \E (K_{n,J(\tau+m')-1} - \tau)
    + 3\delta \E\Big(\sum_{j=J(\tau)}^{J(\tau+m')-1} M_{nj}^2/n^2 \Big) \\
  &\quad+ 3(\delta'/n) \E\Big( \Big(\sum_{t \in I_n(u, K_{n,J(\tau)} - \tau)}
  (\mu_{nt} - \bar \mu_n)\Big)^2 +
  \sum_{j=J(\tau)+1}^{J(\tau + m')-1}\Big( \sum_{t \in I_n(u,M_{nj})}
  (\mu_{nt} - \bar{\mu}_n) \Big)^2 \Big) \\
  & < C_1 \delta + C_2 p_n \delta'
  \end{align*}
  for some finite constants $C_1$ and $C_2$.
  The second inequality comes from writing
  \begin{align*}
    Z_n(\tau, m) &= Z_n'(\tau, m)
    + \tfrac{1}{\sqrt{n}} \sum_{t \in I_n(\tau, m)}
    \big((\mu_{nt} - \bar \mu_n) + (\bar \mu_n - \bar X_n)\big) \\
    &= Z_n'(\tau, m)
    + \tfrac{1}{\sqrt{n}} \sum_{t \in I_n(\tau, m)} (\mu_{nt} - \bar \mu_n)
    - (m/n) Z'_n(\tau, n),
  \end{align*}
  the third follows by construction of $A$
  and fourth follows from an argument similar to Lemma~\ref{L3}.

>>>>

To see when Theorem~\ref{T1}'s Assumption~2 is satisfied, consider two
other parametrizations of this DGP. First, let $\Delta_n = \Delta_0$
and $T_n = T_0$ for some finite $T_0$. This mimics a setting where a
researcher intends to apply the bootstrap to a post-break sample, but
mistakenly includes a small number of pre-break observations as
well. In this case, Assumption 2 holds as long as
$M_n = o(\sqrt{n})$.  Similarly, suppose
$\Delta_n = \Delta / \sqrt{n}$ and $T_n = \pi n$, which mimics a
setting with a moderate break with unknown timing.  Then
$D_n \approx \Delta c (1 - c)$ and Assumption 2 again holds as long as
$M_n = o(\sqrt{n})$. Consequently, using block sizes of length
$o(\sqrt{n})$ ensures that Theorem 1's assumptions on the dispersion
of the $\mu_{nt}$ hold when there is a moderate amount of
heterogeneity---either because the sample is chosen to lie after a
large break estimated break or because any breaks are smaller in
magnitude and, consequently, potentially difficult to detect. When the
degree of heterogeneity is very small or nonexistent, so
$\sum_t (\mu_{nt} - \bar \mu_n)^2$ grows even slower, then larger
block lengths can also satisfy this condition.

@@

\begin{lem}\label{L3}
Under the conditions of Theorem~\ref{T1},
\begin{equation}\label{eq:40}
  \E \Big(\sum_{j=1}^{J_n} \Big( \sum_{t = K_{n,j-1}+1}^{K_{nj}}
  (\mu_{nt}^{*} - \bar \mu_n)\Big)^2\Big) = o(n p_n)
\end{equation}
and $\bar \mu_n^{*} = \E^{*} \bar \mu_n^{*} + o_{L_2}(n^{-1/2})$.
\end{lem}
\begin{proof}
  The second result, on $\bar\mu_n^*$, is an immediate implication of
  Equation~\eqref{eq:40}. To show~\eqref{eq:40}, observe that
  \begin{align*}
    \E \Big(\sum_{j=1}^{J_n} &\Big( \sum_{t= K_{n,j-1} +1}^{K_{nj}}
    (\mu_{nt}^{*} - \bar \mu_n)\Big)^2\Big)= \\
    &= \E\Big(\sum_{j=1}^{J_n} \big\lvert \sum_{s,t = K_{n,j-1}+1}^{K_{nj}}
    \E \big((\mu_{ns}^{*} - \bar{\mu}_n)
    (\mu_{nt}^{*} - \bar{\mu}_n) \mid \Ms \big) \big\rvert \Big)\\
    &\leq \E\Big( \sum_{j=1}^{J_n}
    \sum_{k,l = -n}^{n} \tfrac{1}{n} \sum_{\tau=0}^{n-1} \big\lvert
    (\mu_{n,\tau+k} - \bar{\mu}_n) (\mu_{n,\tau+l} - \bar{\mu}_n) \;
    1\{ \tau + k, \tau + l \in I_n(\tau, M_{nj})\} \big\rvert \Big)\\
    &\leq \E\Big(\sum_{j=1}^{J_n} \Big[
    \sum_{k = -n}^{n} \big(\tfrac{1}{n} \sum_{\tau=0}^{n-1}
    (\mu_{n,\tau+k} - \bar{\mu}_n)^2
    1\{ \tau + k \in I_n(\tau, M_{nj})\}\big)^{1/2}\Big]^{2}\Big)\\
    &= \E\Big(\sum_{j=1}^{J_n}
    M_{nj}^2 \Big\lvert \tfrac{1}{n} \sum_{\tau=0}^{n-1}
    (\mu_{n,\tau} - \bar{\mu}_n)^2 \Big\lvert\Big)\\
    &= o(n p_n).
  \end{align*}
  The first equality follows from the LIE. The first inequality
  exploits the fact that the conditional expectation in the previous
  expression averages over the block's starting period but treats
  every other random variable as contstant. (The argument of the
  indicator function and the indices of the summations are
  deliberately chosen to accomodate blocks that include observations
  from both the beginning and the end of the series.) The second
  inequality is an application of the Cauchy-Schwarz inequality to the
  inner-most summation. The next equality holds because the
  summation
  \[
    \sum_{\tau=0}^{n-1} (\mu_{n,\tau+k} - \bar{\mu}_n)^2
    1\{ \tau + k \in I_n(\tau, M_{nj})\}
  \]
  is zero for all but $M_{nj}$ values of $k$ and is a nonzero constant
  for those $k$. The last line holds by Lemma~\ref{L1} and assumption.
\end{proof}

@@

\begin{proof}
  We will only present a proof of~\eqref{eq:30}. Uniform integrability
  of the first family of random variables is simply a restatement of
  that expression and uniform integrability of $Z_{nj}^{*2} n / M_{nj}$
  follows an abbreviated version of the proof of~\eqref{eq:30}.

@@

  Since $Z_{nj}^*$ is a martingale difference array, Doob's maximal
  inequality for martingales \citep[see][Theorem 15.15, for
  example]{Dav:94} along with the construction of $Z_{nj}^{*}$ gives
  a bound for the first term on the \allcaps{RHS} of~\eqref{eq:6},
  \begin{align*}
  \E \Em^* & \Big[\ind(A) \tfrac{n}{m'}\max_{m = 1,\dots, m'}
  \Big(Z_n^*(\tau, K_{n,J(\tau)} - \tau)
  + \sum_{j=J(\tau)+1}^{J(\tau + m)-1} Z_{nj}^*\Big)^2\Big] \\
  &\leq \tfrac{1}{m'} \sum_{u = 0}^{n-1} \E \Big[\Em \big[
  \ind(A) Z_n(u, K_{n,J(\tau)} - \tau)^2\big] +
  \sum_{j=J(\tau)+1}^{J(\tau + m')-1} \Em \big[\ind(A) Z_n(u, M_{nj})^2]\Big] \\
  &\leq \tfrac{3}{m'} \sum_{u = 0}^{n-1} \E\Big[\Em\big[
  \ind(A) Z'_n(u, K_{n,J(\tau)} - \tau)^2\big]
  + \tfrac{1}{n} \Pm[A] \Big(\sum_{t \in I_n(u, K_{n,J(\tau)} - \tau)}
  (\mu_{nt} - \bar \mu_n)\Big)^2 \\
  & \quad+ \sum_{j=J(\tau)+1}^{J(\tau + m')-1}
  \Em \big[\ind(A) Z_n'(u, M_{nj})^2]
  + \tfrac{1}{n} \Pm[A]
  \sum_{j=J(\tau)+1}^{J(\tau + m')-1}\Big( \sum_{t \in I_n(u,M_{nj})}
  (\mu_{nt} - \bar{\mu}_n) \Big)^2 \\
  & \quad+ \Em\big( \ind(A) Z_n'(u, n)^2 \big)
  \sum_{j=J(\tau)}^{J(\tau+m')-1} M_{nj}^2/n^2 \Big] \\
  &\leq (3\delta/m') \E (K_{n,J(\tau+m')-1} - \tau)
    + 3\delta \E\Big(\sum_{j=J(\tau)}^{J(\tau+m')-1} M_{nj}^2/n^2 \Big) \\
  &\quad+ 3(\delta'/n) \E\Big( \Big(\sum_{t \in I_n(u, K_{n,J(\tau)} - \tau)}
  (\mu_{nt} - \bar \mu_n)\Big)^2 +
  \sum_{j=J(\tau)+1}^{J(\tau + m')-1}\Big( \sum_{t \in I_n(u,M_{nj})}
  (\mu_{nt} - \bar{\mu}_n) \Big)^2 \Big) \\
  & < C_1 \delta + C_2 p_n \delta'
  \end{align*}
  for some finite constants $C_1$ and $C_2$.
  The second inequality comes from writing
  \begin{align*}
    Z_n(\tau, m) &= Z_n'(\tau, m)
    + \tfrac{1}{\sqrt{n}} \sum_{t \in I_n(\tau, m)}
    \big((\mu_{nt} - \bar \mu_n) + (\bar \mu_n - \bar X_n)\big) \\
    &= Z_n'(\tau, m)
    + \tfrac{1}{\sqrt{n}} \sum_{t \in I_n(\tau, m)} (\mu_{nt} - \bar \mu_n)
    - (m/n) Z'_n(\tau, n),
  \end{align*}
  the third follows by construction of $A$
  and the fourth follows from an argument similar to Lemma~\ref{L3}.
